---
title: "simu1_stat_far_noAUC"
author: "Beige Jin"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(root.dir = "/Users/bei/BeiWorld/Research/2022CMRIA/CMR_IA/Analysis/simu1_recog_recsim")
```

### Set Up

```{r lib}
library(MASS)
library(lme4)
library(RePsychLing)
library(lmerTest)
library(car)
library(expm)
getwd()
```

```{r data}
df = read.csv("../simu1_data/simu1_David_hrfar.csv")
df$log_lag_disp = factor(df$log_lag_disp, levels = c(1, 8, 21, 55, 149))
df$roll_cat_len_level = factor(df$roll_cat_len_level, levels = c("0-1", ">=2"))
```

```{r func}
format.random.effect.zc = function(random.effect, group="subject_ID"){
  return(paste("(", random.effect, "||", group, ")"))
}

format.random.effect.wc = function(random.effect, group="subject_ID"){
  return(paste("(", random.effect, "|", group, ")"))
}
```

### Full LMM

If we fit the full LMM, there is too many parameters.

```{r}
# lmer(hr ~ log_lag_disp * roll_cat_len_level + (log_lag_disp * roll_cat_len_level | subject_ID), data=df)
```

### Zero-corr LMM

Now we try zero-correlation LMM. With all factors, it is singular.

```{r}
full.fixed.effect = "log_lag_disp * roll_cat_len_level"
full.random.effect = "(dummy(log_lag_disp, \"8\") + dummy(log_lag_disp, \"21\") + dummy(log_lag_disp, \"55\") + dummy(log_lag_disp, \"149\")) * dummy(roll_cat_len_level, \">=2\")"

formula.zc.full = paste("far ~", full.fixed.effect, "+", format.random.effect.zc(full.random.effect))
cat(formula.zc.full)
mod.zc.full = lmer(formula.zc.full, data=df)
summary(mod.zc.full)
summary(rePCA(mod.zc.full))
```

### Zero-corr LMM -2

Eliminate those with near zero variance. Not singular!

```{r}
random.effect.todrop = "(dummy(log_lag_disp, \"8\") + dummy(log_lag_disp, \"21\"))"
reduced.random.effect = paste(full.random.effect, "-", random.effect.todrop)

formula.zc.drop2 = paste("far ~", full.fixed.effect, "+", format.random.effect.zc(reduced.random.effect))
cat(formula.zc.drop2)
mod.zc.drop2 = lmer(formula.zc.drop2, data=df)
summary(mod.zc.drop2)
```

### LMM -2 full corr

In the following we want to find a maximal correlation matrix.

Extending the reduced LMM with all correlation parameters. It refuses to converge.

```{r}
random.effect.todrop = "(dummy(log_lag_disp, \"8\") + dummy(log_lag_disp, \"21\"))"
reduced.random.effect = paste(full.random.effect, "-", random.effect.todrop)

formula.wc.drop2.full = paste("far ~", full.fixed.effect, "+", format.random.effect.wc(reduced.random.effect))
cat(formula.wc.drop2.full)
mod.wc.drop2.full = lmer(formula.wc.drop2.full, data=df)

# it can't converge even with replacing start values
# theta <- getME(mod.wc.drop2.full, "theta")
# start_vals <- list(theta = theta)
# mod.wc.drop2.full = lmer(formula.wc.drop2.full, data=df, start = start_vals)

summary(mod.wc.drop2.full)
attr(VarCorr(mod.wc.drop2.full)[["subject_ID"]], "correlation")
```

### LMM -2 corr -1

Cut out `dummy(log_lag_disp, "149"):dummy(roll_cat_len_level, ">=2")` whose correlations are most near zero. Not converge.

```{r}
random.effect.todrop = "(dummy(log_lag_disp, \"8\") + dummy(log_lag_disp, \"21\") + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\"))"
reduced.random.effect = paste(full.random.effect, "-", random.effect.todrop)
random.effect2 = "0 + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\")"

formula.wc.drop2.cut1 = paste("far ~", full.fixed.effect, "+", format.random.effect.wc(reduced.random.effect), "+", format.random.effect.wc(random.effect2))
cat(formula.wc.drop2.cut1)
mod.wc.drop2.cut1 = lmer(formula.wc.drop2.cut1, data=df)

# it can't converge even with replacing start values
# theta <- getME(mod.wc.drop2.cut1, "theta")
# start_vals <- list(theta = theta)
# mod.wc.drop2.cut1 = lmer(formula.wc.drop2.cut1, data=df, start = start_vals)

summary(mod.wc.drop2.cut1)
Matrix::bdiag(sapply(VarCorr(mod.wc.drop2.cut1), attr, "correlation"))
```

### LMM -2 corr -2

Cut out `dummy(log_lag_disp, "8"):dummy(roll_cat_len_level, ">=2")` whose correlations are most near zero. Singular.

```{r}
random.effect.todrop = "(dummy(log_lag_disp, \"8\") + dummy(log_lag_disp, \"21\") + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\") + dummy(log_lag_disp, \"8\"):dummy(roll_cat_len_level, \">=2\"))"
reduced.random.effect = paste(full.random.effect, "-", random.effect.todrop)
random.effect2 = "0 + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\")"
random.effect3 = "0 + dummy(log_lag_disp, \"8\"):dummy(roll_cat_len_level, \">=2\")"

formula.wc.drop2.cut2 = paste("far ~", full.fixed.effect, "+", format.random.effect.wc(reduced.random.effect), "+", format.random.effect.wc(random.effect2), "+", format.random.effect.wc(random.effect3))
cat(formula.wc.drop2.cut2)
mod.wc.drop2.cut2 = lmer(formula.wc.drop2.cut2, data=df)
summary(mod.wc.drop2.cut2)
Matrix::bdiag(sapply(VarCorr(mod.wc.drop2.cut2), attr, "correlation"))
```

### LMM -2 corr -3

Cut out `dummy(roll_cat_len_level, ">=2")` whose correlations are most near zero. Not singular!!

```{r}
random.effect.todrop = "(dummy(log_lag_disp, \"8\") + dummy(log_lag_disp, \"21\") + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\") + dummy(log_lag_disp, \"8\"):dummy(roll_cat_len_level, \">=2\") + dummy(roll_cat_len_level, \">=2\"))"
reduced.random.effect = paste(full.random.effect, "-", random.effect.todrop)
random.effect2 = "0 + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\")"
random.effect3 = "0 + dummy(log_lag_disp, \"8\"):dummy(roll_cat_len_level, \">=2\")"
random.effect4 = "0 + dummy(roll_cat_len_level, \">=2\")"

formula.wc.drop2.cut3 = paste("far ~", full.fixed.effect, "+", format.random.effect.wc(reduced.random.effect), "+", format.random.effect.wc(random.effect2), "+", format.random.effect.wc(random.effect3), "+", format.random.effect.wc(random.effect4))
cat(formula.wc.drop2.cut3)
mod.wc.drop2.cut3 = lmer(formula.wc.drop2.cut3, data=df)

# replace start values could make it converge
for (i in 1:4){
  theta <- getME(mod.wc.drop2.cut3, "theta")
  start_vals <- list(theta = theta)
  mod.wc.drop2.cut3 = lmer(formula.wc.drop2.cut3, data=df, start = start_vals)
}

summary(mod.wc.drop2.cut3)
```

Try to combine the three cut out components. Not singular!!

```{r}
random.effect.todrop = "(dummy(log_lag_disp, \"8\") + dummy(log_lag_disp, \"21\") + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\") + dummy(log_lag_disp, \"8\"):dummy(roll_cat_len_level, \">=2\") + dummy(roll_cat_len_level, \">=2\"))"
reduced.random.effect = paste(full.random.effect, "-", random.effect.todrop)
random.effect2 = "0 + dummy(log_lag_disp, \"149\"):dummy(roll_cat_len_level, \">=2\") + dummy(log_lag_disp, \"8\"):dummy(roll_cat_len_level, \">=2\") + dummy(roll_cat_len_level, \">=2\")"

formula.wc.drop2.cut3.comb = paste("far ~", full.fixed.effect, "+", format.random.effect.wc(reduced.random.effect), "+", format.random.effect.wc(random.effect2))
cat(formula.wc.drop2.cut3.comb)
mod.wc.drop2.cut3.comb = lmer(formula.wc.drop2.cut3.comb, data=df)
summary(mod.wc.drop2.cut3.comb)
```

### Summary

In summary, we will adopt mod.wc.drop2.cut3.comb.

```{r}
summary(mod.wc.drop2.cut3.comb)
anova(mod.wc.drop2.cut3.comb)
```

